---
title: "Proyecto 1"
author: "Camilo Rojas"
date: "2023-10-11"
output: html_document
---


```{r, echo=TRUE}
library(knitr)
opts_chunk$set(echo = TRUE)

```

First question
==============

I load the dataset and named it as ***data***

```{r}
library(rio)

data <- import("./activity.csv")

data$steps <- as.numeric(data$steps)
data$date <- as.Date(data$date)
```

second question
=================


```{r}
total_steps <- sum(data$steps, na.rm = T)
```

1. the number total of steps in this dataset is `r total_steps`

2. i did the histogram per day vs number of steps
```{r, fig.width=8, fig.height=4}
library(dplyr)
library(ggplot2)

data %>%
 filter(!is.na(steps)) %>%
 ggplot() +
  aes(x = date, weight = steps) +
  geom_bar(fill = "#112446") +
  labs(
    x = "Day",
    y = "number of steps",
    title = "steps per day"
  ) +
  theme_classic()

```
3. the next table show report the mean and median of the total number of steps taken per day

```{r, results= 'asis'}
library(kableExtra)
data1 <- data %>% 
        select(date, steps) %>% 
        group_by(date) %>% 
        summarise(median = median(steps,na.rm = T), mean = mean(steps,na.rm = T))
data1$mean <- ifelse(is.nan(data1$mean),0,data1$mean)
data1$median <- ifelse(is.nan(data1$median),0,data1$median)
kbl(data1) %>% 
        kable_styling()

```
third question
=================

time series plot for 5-minute interval

```{r}
interval5 <- data %>% 
        filter(interval == 5)

interval5 %>%
 filter(!is.na(steps)) %>%
 ggplot() +
  aes(x = date, weight = steps) +
  geom_bar(fill = "#112446") +
  theme_minimal()
```

the day that has the maximun number of steps is 2012-10-10

fourd question
=================

```{r}

steps_na <- sum(is.na(data$steps))
date_na <- sum(is.na(data$date))
interval_na <- sum(is.na(data$interval))
```

1. In this dataset i could found that we has `r steps_na` NA values at the steps column, `r date_na` NA values at the date column and `r interval_na` NA values at the interval column 

2. fill de Na values in the dataset

```{r}

fill_data <- merge(data, data1, by = "date")

fill_data$steps_adj <- ifelse(is.na(fill_data$steps), fill_data$mean, fill_data$steps)
        

fill_data %>%
 ggplot() +
  aes(x = date, weight = steps) +
  geom_bar(fill = "#112446") +
  theme_minimal()
```

when all the values filled we can see that the values for the mean and median per day looks like 

```{r}
data2 <- fill_data %>% 
        select(date, steps) %>% 
        group_by(date) %>% 
        summarise(median = median(steps), mean = mean(steps))

kbl(data2) %>% 
        kable_styling()
```